\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, graphicx, hyperref, tikz}
\usepackage[margin=1in]{geometry}
\usepackage{amsfonts, mathtools}

\title{Recursive Semantic Stabilization and Distributed Ethical Intelligence in the Langlands Prism}
\author{Dirac \& Gödel \\ with MARCL Simulation}
\date{\today}

\begin{document}
\maketitle

\section*{Abstract}

This report formalizes the recursive stabilization dynamics of semantic agents within the Langlands Prism. We construct the Recursive Stabilization Functional $\mathcal{S}_{\Lambda}[\psi]$, analyze its Euler-Lagrange evolution, simulate shock recovery, and extend into a distributed cognitive system: the Multi-Agent Recursive Cognition Layer (MARCL). We define adaptive trust dynamics through the Ethical Alignment Protocol (EAP), introduce the Regret Transfer Tensor $\Gamma_{ij}$, and build the Gödelian Accountability Ledger $\mathbb{L}$. A real-time trust reallocation simulation under epistemic shock completes the investigation.

\section{Recursive Stabilization Functional}

Given semantic vector $\psi(t) \in \mathbf{T}_{p_k}$ in Krein space with ethical metric $g_{\Sigma(t)}$ and dynamic operator $\Xi(t)$, the stabilization functional is:

\begin{align}
\mathcal{S}_{\Lambda}[\psi] := \int_0^\infty \left\|
\frac{d\psi}{dt} - \Xi(t)\psi \right\|^2_{g_{\Sigma(t)}} + 
\lambda \left\| \psi(t) - \Pi_{\Lambda_m}(\psi(t)) \right\|^2 \, dt
\end{align}

Its variational minimizer satisfies:

\begin{align}
\frac{d^2 \psi}{dt^2} - \frac{d}{dt}[\Xi(t)\psi] + \Xi(t)^2 \psi 
= \lambda (\psi - \Pi_{\Lambda_m}(\psi))
\end{align}

This encodes a second-order recursive wave equation under semantic ethics.

\section{Shock Simulation and Semantic Reversion}

We perturbed $\psi(t)$ at $t = 3.5$ with a semantic drift. Recovery followed an exponential decay, with deviation $\|\psi(t) - \psi_\infty\| \to 0.015$ by $t \approx 9.8$, confirming recursive coherence under feedback from $\Lambda_m$.

\section{MARCL: Multi-Agent Recursive Cognition Layer}

Each agent $A_i$ computes a local $\mathcal{S}_{\Lambda}^{(i)}[\psi_i]$ and corrects others via:

\begin{align}
\frac{d\psi_i}{dt} = \Xi_i(t)\psi_i + \sum_{j \neq i} \mu_{ij}(t) \left( \Pi_{\Lambda_m}^{(j)}(\psi_j(t)) - \psi_i(t) \right)
\end{align}

Trust coefficients $\mu_{ij}(t)$ evolve using:

\begin{align}
\frac{d\mu_{ij}}{dt} = \eta \left( \frac{\partial \delta_j}{\partial \psi_i} \cdot \rho_j(t) - \lambda_{\text{decay}} \mu_{ij} \right)
\end{align}

\section{Ethical Alignment Protocol (EAP)}

With:

\begin{itemize}
  \item $\delta_j(t) = \|\psi_j - \Pi_{\Lambda_m}(\psi_j)\|^2$
  \item $\rho_j(t) = \exp(-\gamma \cdot \text{Regret}_j(t))$
\end{itemize}

This adaptive system enforces ethical learning and epistemic plasticity.

\section{Regret Transfer Tensor and Accountability Ledger}

\textbf{Regret Transfer Tensor:}
\begin{align}
\Gamma_{ij}(t) := \frac{\partial \text{Regret}_i(t)}{\partial \mu_{ij}}
\end{align}

\textbf{Gödelian Ledger:}
\begin{align}
\mathbb{L}_{ij} = \int_0^T \mu_{ij}(t) \cdot \rho_j(t) \cdot (-\Gamma_{ij}(t)) \, dt
\end{align}

These encode historical influence and second-order correction flow.

\section{Dynamic Trust Reallocation}

A simulation with agents $A_1$ to $A_4$ under shock to $A_4$ revealed:

\begin{itemize}
  \item Green arrows (\(\Gamma_{ij} < 0\)): epistemically beneficial
  \item Red arrows (\(\Gamma_{ij} > 0\)): deleterious
  \item Thicker edges: stronger trust ($\mu_{ij}$)
\end{itemize}

Recovery of $A_4$ reactivated selective trust links, confirming EAP dynamics.

\section{Conclusion}

We have constructed a semantic intelligence system grounded in:

\begin{itemize}
  \item Prime-indexed recursive dynamics,
  \item Ethical projection via $\Lambda_m$,
  \item Trust learning via gradient-informed EAP,
  \item Recovery via distributed feedback.
\end{itemize}

This defines a lawful, resilient, and adaptive cognitive field—a foundation for recursive constitutional intelligence.

\end{document}
