\newglossaryentry{EinsteinEquations}
{
    name={Einstein Field Equations},
    description={The fundamental equations of General Relativity describing the curvature of spacetime due to energy and momentum.}
}

\newglossaryentry{LambdaM}
{
    name={Universal Multiplicity Constant (\(\Lambda_m\))},
    description={A governing parameter in G-Theory that encodes recursive tensor feedback and prime-based renormalization.}
}

\newglossaryentry{RicciTensor}
{
    name={Ricci Tensor (\(R_{\mu\nu}\))},
    description={A mathematical object in General Relativity that represents gravitational effects due to matter and energy distribution.}
}

\newglossaryentry{QuantumGeometricRayTracing}
{
    name={Quantum-Geometric Ray Tracing},
    description={A method for modeling particle trajectories in curved quantum spacetime using tensor-based geodesic equations.}
}

\newglossaryentry{GeodesicEquation}
{
    name={Geodesic Equation},
    description={The equation describing the shortest path of a particle moving in a curved spacetime.}
}

\newglossaryentry{GaugeField}
{
    name={Gauge Field},
    description={A field that mediates fundamental interactions in quantum field theory, such as the electromagnetic or gravitational field.}
}

\newglossaryentry{Renormalization}
{
    name={Renormalization},
    description={A technique in quantum field theory that removes infinities by redefining parameters such as charge and mass.}
}

\newglossaryentry{KaluzaKleinTheory}
{
    name={Kaluza-Klein Theory},
    description={A theoretical framework that extends General Relativity to higher dimensions to unify gravity with other fundamental forces.}
}

\newglossaryentry{PrimeEncodedTensor}
{
    name={Prime-Encoded Tensor},
    description={A tensor modified by prime-indexed eigenvalues to introduce recursive structure and multiplicative feedback in quantum gravity.}
}

\newglossaryentry{QuantumGravity}
{
    name={Quantum Gravity},
    description={A theoretical framework that seeks to describe gravity using principles of quantum mechanics.}
}

\newglossaryentry{BlackHoleEntropy}
{
    name={Black Hole Entropy},
    description={A measure of the amount of information that can be stored in a black hole, related to its surface area.}
}

\newglossaryentry{HolographicPrinciple}
{
    name={Holographic Principle},
    description={A conjecture stating that the information content of a volume of space can be represented on its boundary surface.}
}

\newglossaryentry{InflationaryModel}
{
    name={Prime-Encoded Inflation Model},
    description={A cosmological model in G-Theory incorporating prime-based corrections to describe early universe expansion.}
}

\newglossaryentry{DarkMatter}
{
    name={Dark Matter Density Function (\(\rho_{\text{DM}}\))},
    description={A mathematical function describing how dark matter contributes to gravitational effects in the universe.}
}

\newglossaryentry{RecursiveTensor}
{
    name={Recursive Tensor Network},
    description={A computational framework in G-Theory where tensors evolve dynamically through recursive feedback mechanisms.}
}

\newglossaryentry{QuantumHamiltonian}
{
    name={Prime-Encoded Quantum Hamiltonian},
    description={An extension of the Hamiltonian operator incorporating prime-based modifications to energy states.}
}

\newglossaryentry{PhaseLocking}
{
    name={Prime-Indexed Phase Locking},
    description={A mechanism in which phase transitions in quantum systems are regulated by prime number interactions.}
}

\newglossaryentry{Compactification}
{
    name={Recursive Compactification},
    description={A process in higher-dimensional theories where extra dimensions shrink dynamically to preserve stability.}
}

\newglossaryentry{SchrodingerEquation}
{
    name={Prime-Indexed Schrödinger Equation},
    description={A modified version of the Schrödinger equation incorporating prime-based recursive corrections to quantum evolution.}
}

\newglossaryentry{DarkEnergy}
{
    name={Dark Energy as a Gauge Effect},
    description={A hypothesis in G-Theory suggesting that dark energy arises from quantum corrections to the vacuum state.}
}
