\appendix
\section{Reproducibility Capsule: Self-Verifying End-to-End Trust Chain}
\label{app:capsule}

This appendix describes the self-verifying reproducibility capsule accompanying the Universal Atomic Calculator (UAC) framework and its integration with Multiplicity Social Physics (MSP). The capsule enables independent verification of claims from neutral-atom hardware readouts through Rust-mediated facts to Lean 4 machine-checked invariants.

\subsection{Capsule Architecture}
The capsule is distributed as a single archive containing:
\begin{itemize}
    \item \texttt{paper.pdf} -- Anchoring narrative with citations to sealed facts.
    \item \texttt{data/} -- Signed raw Qiskit/Pasqal outputs (\texttt{h2\_qiskit\_log.json}) and certified \texttt{ReadOnlySignatureFact} stream.
    \item \texttt{rust-bridge/} -- Sedona Spine-compliant immutable pipeline (sole source of truth).
    \item \texttt{lean-proofs/} -- Zero-\texttt{sorry}, minimal-core formalization.
    \item \texttt{harness.sh} -- One-command verification orchestrator.
\end{itemize}

Full file tree and build instructions are in the capsule \texttt{README.md}.

\subsection{Hardware-to-Fact Pipeline (Rust Layer)}
Raw neutral-atom measurement logs are processed by the Rust bridge, which:
\begin{enumerate}
    \item Verifies cryptographic signatures.
    \item Applies L0-bounded filtering (hard \texttt{SIG\_GOV\_KILL} on breach).
    \item Emits sealed \texttt{ReadOnlySignatureFact} objects.
\end{enumerate}

This enforces the Path of Integrity: \texttt{Hardware} $\to$ \texttt{Engine (Rust)} $\to$ \texttt{SDK} $\to$ higher layers. Agents/UI may only transform, never override.

\subsection{Formal Verification (Lean 4 Layer)}
The Lean 4 module proves key UAC invariants under the assumption of faithful hardware representation:

\begin{lstlisting}[language=Lean]
-- UAC_Invariants.lean excerpt
theorem uac_l0_ceiling (facts : List QiskitReadOnlyFact) :
  ∀ f ∈ facts, f.simulated_norm ≤ 3900 := by
    -- decide reduction; zero sorry
    simp [List.all, decide]
    -- proof proceeds via bounded operator norms from UAC theory
\end{lstlisting}

Additional theorems cover resonance coherence (linking to MSP embodied viability $\mathcal{E}(r,t)$) and convergence under decoherence noise.

\subsection{UAC-MSP Integration}
The capsule grounds UAC qudit/neutral-atom self-simulation in MSP observables. The Socio-Atomic Model and indexed social factors remain **heuristic/pedagogical analogies** (see \S\ref{sec:scope}).

\begin{table}[H]
\centering
\caption{UAC Hardware Facts $\to$ MSP Observables (Heuristic Mapping)}
\begin{tabular}{ll}
\toprule
UAC Primitive & MSP Observable (Heuristic) \\
\midrule
Qudit multiplicity / Rydberg entanglement & Multiplicity formula $2R + 1 = M$ \\
Decoherence variance bounds & Resonance coherence $\mathcal{R}(r,t)$ \\
Chemical accuracy benchmarks & Embodied viability $\mathcal{E}(r,t)$ / MSC valuation input \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Scope and Limitations:} The mappings are pedagogical scaffolding and do not constitute formal isomorphisms or advances to the mathematical core. All formal claims are confined to the Lean 4 proofs. \socialdisclaimer

\subsection{Verification Instructions}
Run \texttt{./harness.sh} after unpacking. Expected output confirms all chains intact. The harness also supports controlled fault-injection replay (opt-in) to demonstrate fail-closed behavior.

This capsule satisfies ADR requirements for machine-checked governance and provides reviewers with executable trust.

\subsection{Defense and Dissemination}
Placed in public domain (core math/proofs) under CC BY-NC-SA 4.0. Commercial UAC/MSP implementations require separate licensing via Citizen Gardens.
