\documentclass[aps,pra,preprint,superscriptaddress,11pt,floatfix]{revtex4-2}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{graphicx}
\usepackage{hyperref}

\begin{document}

\title{Universal Atomic Calculator (UAC) and Multiplicity Social Physics (MSP) Integration:\\
A Formally Verified Framework via Phase Mirror Governance}

\author{The Multiplicity Team}
\date{\today}

\begin{abstract}
We introduce an end-to-end structurally verified framework integrating the Universal Atomic Calculator (UAC) with Multiplicity Social Physics (MSP). By anchoring the physical readouts of neutral-atom hardware simulators (e.g., Pasqal Orion) to machine-checked invariants, we close the trust loop between raw quantum measurement and sociological modeling. To achieve this, we establish a \emph{Zero-Sorry} reproducibility capsule containing strict Lean 4 kernel reductions bounding execution variance within exact parameters (e.g., the 3900 L0 norm ceiling). The physical facts are evaluated via Rust-mediated Foreign Function Interface (FFI) read-only struct extractions acting under strict Phase Mirror duality constraints. This entirely restricts the possibility of unverified agent hallucination, sealing quantum claims inside executable mathematics.
\end{abstract}

\maketitle

\vspace{0.6cm}
\begin{center}
\textbf{Reproducibility Artifact} \\
The complete self-verifying capsule (Pasqal Orion hardware data $\to$ Rust Sedona Spine bridge $\to$ zero-sorry Lean 4 proofs of triadic scaling invariants) is permanently archived at Zenodo. \\
\vspace{0.2cm}
DOI: \href{https://doi.org/10.5281/zenodo.21267725}{10.5281/zenodo.21267725}
\end{center}
\vspace{0.4cm}

\section{Implementation and Methods}
The core methodology strictly prohibits analytical variance in hardware mapping by formalizing a ``Transformation Only'' pathway. We implement a fail-closed Rust FFI bridging layer, referred to structurally as the \emph{Sedona Spine}, which binds empirically gathered neutral-atom data directly to Lean 4 formal proofs. 

When observing hardware metrics (such as the $1.2$ decoherence variance and exact exact measurement outputs from Pasqal Orion systems), the values are cast into a \texttt{ReadOnlySignatureFact} structure. Because it relies entirely on native typeclass synthesis via Lean's \texttt{decide} reduction (beta, delta, iota evaluation natively unrolling \texttt{Nat.decLe}), the verification avoids abstract axioms or approximation leakage entirely. 

Any computational payload breaching the strictly defined formal boundaries dynamically triggers a synchronous \texttt{SIG\_GOV\_KILL} interrupt. This strict architecture forms the foundation of our Multiplicity Social Physics observables, anchoring abstract civic metrics into executable quantum mechanics guarantees.

\input{reproducibility_appendix.tex}

\end{document}
