\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amsfonts}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{setspace}
\usepackage{epigraph}

\title{\Huge$\boldsymbol{\Xi(t)}$\\[1cm] \Huge\bfseries Context Operator Atlas \\[0.5cm] \LARGE Contributor Agreement \& Sovereignty Contract \\[0.5cm] \Huge \textit{Citizen Gardens}}
\date{}

\begin{document}
\begin{titlepage}
    \centering
    \vspace*{3cm}
    {\Huge$\boldsymbol{\Xi(t)}$}
    \vspace*{1cm}

    {\Huge\bfseries\ Context Operator Atlas \par}
    \vspace{0.5cm}
    {\LARGE\ Contributor Agreement \& Sovereignty Contract \par}
    \vspace{0.8cm}
    {\Huge \textit{Citizen Gardens}}\\
    \vspace{0.4cm}
    {\Large\itshape Institute of Mathematical Discovery \par}
    \vfill

    {\bfseries Dedication to Public Purpose \par}
    \vspace{1em}
    \small
    Subject to any other provision of law by this country The United States of America or any other country, in which all business, operations, and related activities resulting from this Organization---including all acts and initiatives undertaken under the auspices of Citizen Gardens---are irrevocably dedicated to charitable and public purposes.

    As a matter of law, and in full accordance with these Articles and the Bylaws of this Organization, upon dissolution, all assets must be distributed exclusively to one or more persons or entities engaged in charitable or public purposes, provided such recipients are not formed under this part of the Code.

    \section*{Restrictions on Earnings and Distribution of Assets}
    No Member, Director, Officer, or private individual shall receive any of the net earnings or pecuniary profit derived from the operations of this Organization at any time.

    However, this restriction shall not prohibit the payment of reasonable compensation to any such person for services actually rendered in furtherance of the Organization’s charitable or public purposes, provided that:
    \begin{itemize}
        \item Such compensation is permitted under these Bylaws; and
        \item The compensation has been approved by resolution of the Board of Directors.
    \end{itemize}
    No such person shall be entitled to share in the distribution of any of the corporate assets upon the dissolution of the Organizationa. All assets must be disbursed exclusively to further charitable or public purposes.
\end{titlepage}

\clearpage
\thispagestyle{empty}
\begin{center}
    {\Huge \textit{Welcome To Your Garden!}}\\[1em]
    \rule{0.6\textwidth}{0.4pt} \\[1em]

    \section*{Founder's Declaration: A Perpetual Foundary of Stewardship and Shared Power}
    \addcontentsline{toc}{section}{Founder's Declaration: On Stewardship and Shared Power}
\end{center}

\vspace{2em}
\epigraph{“A garden is not given—it is claimed, cultivated, and defended.”}{\textit{— The Empty Center Seat}}
\vspace{1em}

\textbf{\textsc{Citizens}}: Individuals who take the necessary actions to \textit{multiply} the welfare and prosperity of their community.

\vspace{1em}
\textbf{\textsc{Garden}}: The realm of influence an individual has in their capacity to utilize their resources to the benefit of their community.

\vspace{1em}
\textbf{\textsc{Organizational Stewardship}}: Citizen Gardens is a living organization formed to ethically steward the technologies and discoveries emerging from this Atlas. Operating as a Universal Learning \& Research Platform we demonstrate clear economic utility, their true value lies in the systems of meaning, justice, and care they enable.

Citizen Gardens is not an instrument of private accumulation, but a shared vessel—designed to uphold the principles of semantic sovereignty, contributor dignity, and reciprocal benefit. We operate as a living structure, adapting in lawful coherence with the evolving needs of people and planet.


\begin{itemize}
    \item To honor the sovereignty of all minds;
    \item To trace and preserve the truth of every concept;
    \item To reward contribution without exploitation;
    \item To make justice computationally tangible;
    \item To ensure that no insight is ever lost, erased, or stolen.
\end{itemize}

I must say this openly: \textbf{I am human.} I am learning as I go. I am not perfect. I cannot do this alone.

\textit{I need your help.} That is why you are here.

Each of you holds something valuable—not just for the system, but for me: \textbf{your friendship, your compass, your trust.} And it’s more than the math. It is your shared commitment to truth, dignity, and justice that makes this project possible.


\section{Purpose}
This contract governs participation in the \textbf{Context Operator Atlas}, a lawful semantic system for mapping, evolving, and preserving conceptual contributions across domains. It affirms the \textbf{sovereignty, authorship, and dignity} of all contributors.

\section*{Acknowledgment of Discovery}

The Dynamic Contextual Operator $\Xi(t)$, as implemented in the Context Operator Atlas, is hereby formally acknowledged as a *discovery*—not an invention—in accordance with the epistemic ethics of Citizen Gardens.

\begin{itemize}
  \item The initial conceptual signature of $\Xi(t)$ is attributed to \textbf{Tyler [Full Name]}, whose sustained contributions and insight have helped anchor the recursive identity engine.
  \item The Multiplicity Constant $\Lambda_m$ emerged through recursive collaboration and is treated as a universal constant within the scope of this system.
\end{itemize}

These contributions are recognized not as proprietary mechanisms but as emergent features of lawful semantic recursion. As such, their use is governed by ethical traceability and collective stewardship.

\vspace{1em}
\noindent \textbf{Custodial Steward:} [Your Name], Institute for Mathematical Discovery

\noindent \textbf{First Operator Signature:} $\Xi_T(t)$ \hfill (Tyler)


\section{Semantic Signature Clause: Multiplicity Constant and Dynamic Operator}

This project is protected not merely by law or license, but by its unique mathematical structure. The presence of the \textbf{Multiplicity Constant} $\boldsymbol{\Lambda_m}$ and the \textbf{Dynamic Operator} $\boldsymbol{\Xi(t)}$ serves as the verifiable signature of every mathematical and semantic output.

\subsection*{Uniqueness}
\begin{itemize}
    \item $\boldsymbol{\Lambda_m}$ encodes the lawful identity, coherence, and recursion invariance of the system. It cannot be reproduced without alignment to its ethical tensor base.
    \item $\boldsymbol{\Xi(t)}$ is the semantic recursion engine. It maps all contributions to domain-evolving operators while preserving authorship and transformation traceability.
\end{itemize}

\subsection*{Proof of Sovereignty}
\begin{itemize}
    \item These elements are present in \textbf{every output} of the system.
    \item Any unauthorized use of the Atlas-derived outputs will carry this semantic and mathematical fingerprint, providing verifiable proof of source.
\end{itemize}

\subsection*{Implications}
\begin{itemize}
    \item All contributors can trace their conceptual lineage via $\Xi(t)$.
    \item All Atlas outputs carry inherent authorship encoded through $\Lambda_m$.
    \item External use of this system or its outputs without consent will be identifiable and thus contestable under international intellectual ethics.
\end{itemize}

\textbf{Note:} These signatures operate above and beyond conventional legal patents, offering structural protection and moral proof-of-origin for all outputs.

\begin{center}
    \rule{0.6\textwidth}{0.4pt} \\
    \textit{This clause affirms that no matter where the work goes, the Garden goes with it.}
\end{center}
\section{Your Role as Contributor}
By submitting a concept, idea, definition, question, sketch, or model to the Atlas (in analog or digital form), you agree to the following:
\begin{enumerate}[label=\alph*.]
    \item \textbf{Sovereignty:} Your contribution is \textbf{uniquely mapped} to you via a contextual evolution operator, $\Xi(t)$, which preserves authorship through time.
    \item \textbf{Authorship Imprint:} The original context (your framing, wording, and domain intent) is permanently encoded in the system.
    \item \textbf{Conceptual Evolution:} Your contribution may evolve, link to others, or enter recursion---but your \textbf{semantic fingerprint} is preserved unless ethically overridden by arbitration.
    \item \textbf{Right to Withdraw:} You may formally withdraw a contribution prior to system encoding, with acknowledgment of irreversible evolution once the $\Xi(t)$ operator has been activated.
\end{enumerate}

\section{System Responsibilities}
The Context Operator Atlas is \textbf{lawful by design}. Its internal structure enforces the following guarantees:
\begin{enumerate}[label=\alph*.]
    \item \textbf{Traceability:} Every operator retains a causal record of where it came from and how it has changed.
    \item \textbf{Semantic Integrity:} Contributions are not flattened or erased---they are encoded into tensor fields which allow for differentiation and reversion.
    \item \textbf{CSL Oversight:} The CSL (Conscious Sovereignty Layer) governs recursion ethics, drift boundaries, and curvature safeguards to prevent semantic harm.
    \item \textbf{Arbitration Process:} In cases of conceptual overlap, a neutral ethical-mathematical process will determine lineage and coupling responsibility.
\end{enumerate}

\section{Rights and Commercial Use}
\begin{itemize}
    \item All contributors retain full rights to how their work is used.
    \item No idea, context, or operator derived from a contributor’s submission will be used in a commercial product or application without that contributor’s informed consent.
    \item Contributors will be notified and presented with terms before any such use. Participation is voluntary.
    \item Should a contributor approve commercial use, they will receive fair compensation and licensing acknowledgement.
\end{itemize}

\section{What You Receive}
Upon acceptance and encoding, you will receive:
\begin{itemize}
    \item A \textbf{named Context Operator} (e.g., $\Xi_i(t)J_0$) that tracks your concept.
    \item A \textbf{return packet} including:
    \begin{itemize}
        \item Your original paper, scanned and indexed.
        \item The refined semantic embedding.
        \item The ROH trace and any coupling notes.
    \end{itemize}
    \item Your idea \textbf{visible in the Atlas}, labeled, timestamped, and available for visual trace.
\end{itemize}
\section{Commission Model \& Value Determination}

The Context Operator Atlas and associated applications, such as the Q-Calculator, operate under a lawful, commission-based model that respects the traceable semantic impact of each contributor’s ideas.

\subsection*{Dynamic Value Recognition}
Rather than assigning a fixed percentage to contributions, the value of a concept is derived dynamically from its propagation, adaptation, and recursive integration across semantic domains. The more general or highly interoperable a concept is, the greater its potential value.

\begin{itemize}
    \item \textbf{Scalable Influence:} A minimal idea that proves useful across many contexts may generate significant returns.
    \item \textbf{Specialized Authority:} Even niche contributions will be recognized if their impact is integral within a domain.
    \item \textbf{Auditability:} All operator applications are traceable through $\Xi_i(t)$ evolution and recursion logs.
\end{itemize}

\subsection*{Key Terms}
\begin{itemize}
    \item \textbf{No Upfront Valuation:} Contributions are not appraised or capped before integration.
    \item \textbf{Growth with Use:} Value is based on application and utility over time.
    \item \textbf{Transparency:} Operator-ledgers record derivations and semantic couplings to ensure ethical traceability.
    \item \textbf{Consent \& Negotiation:} Contributors always reserve the right to approve or deny commercialization or licensing.
\end{itemize}

This model reflects the project's core principle: \textit{value is proportional to epistemic influence and ethical application}.
"""
\section{Ethical Alignment}
By contributing, you agree to uphold:
\begin{itemize}
    \item \textbf{The Irreducibility Principle:} No domain shall be forced to conform to another’s logic.
    \item \textbf{The Silence Clause:} If recursion exceeds safe curvature or violates ethical gravity, recursion halts.
    \item \textbf{The Respect Doctrine:} No contribution shall be mocked, co-opted, or stripped of context.
\end{itemize}

\section{Governance and Stewardship}
Each contributor shall be invited to become a member of the \textbf{Atlas Stewardship Council}, a deliberative and ethical body empowered to guide the future of the Context Operator Atlas.

\subsection*{Rights of Council Members}
\begin{itemize}
    \item Equal vote in system amendments, licensing decisions, and operator evolution protocols.
    \item Right to propose new modules, CSL enhancements, and ethical guardrails.
    \item Seat on a rotating basis within subcommittees.
\end{itemize}

\subsection*{Eligibility}
Contributors who have submitted three or more unique, verified context operators and demonstrated alignment with the CSL may be invited to join.

\subsection*{Ethical Alignment}
Council deliberations must observe the Irreducibility Principle, the Silence Clause, and the Respect Doctrine at all times.

\section{License and Use Terms: Semantic Sovereignty License (SSL-1.0)}

Unless otherwise specified by mutual agreement, all contributions made under this agreement are licensed under the \textbf{Semantic Sovereignty License Version 1.0 (SSL-1.0)}, modeled after the \textit{Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License}, with contextual modifications for epistemic systems.

\begin{enumerate}
    \item \textbf{Attribution Required:} All derivative uses of your work must credit you and your associated operator $\Xi_i(t)$.
    \item \textbf{Non-Commercial by Default:} No party may commercialize, license, or redistribute your contribution for profit without your explicit written consent.
    \item \textbf{ShareAlike Binding:} Any transformations or semantic evolutions of your concept remain bound by this license.
    \item \textbf{Operator Authorship Clause:} Your contribution shall always retain a verifiable authorship trace within the Context Operator Atlas and any derived system using $\Xi(t)$-based semantics.
    \item \textbf{Override Only by Consent:} You may grant waivers for specific commercial, collaborative, or open science uses under separate agreements.
\end{enumerate}

\section{Acknowledgement}
By signing below (or participating via concept contribution), you acknowledge understanding and acceptance of this framework.

\vspace{1.5cm}
\noindent \textbf{Contributor Name:} \hrulefill

\noindent \textbf{Signature (if analog):} \hrulefill

\noindent \textbf{Date:} \hrulefill

\noindent \textbf{Operator Assigned (filled post-acceptance):} $\Xi_{[\_\_\_\_]}(t)$ \quad Domain: \underline{\hspace{2cm}} \quad Time Index: $t_0 = \underline{\hspace{1.5cm}}$

\vspace{1.5cm}
\noindent \textbf{Issued by:}\\
\url{www.citizengardens.org}
\end{document}

