\documentclass{article}
\usepackage{amsmath, amssymb, listings}
\title{RESONANCE-CORE: Quantum Monad Formalism}
\author{Ahmad Ali Parr (Architect) \\ SnapKitty Sovereign OS --- SEIT NGO}
\date{2026}

\begin{document}
\maketitle

\section{Quantum Superposition}
A superposition $\mathcal{S}$ is a list of amplitude-value pairs:
\begin{equation}
  \mathcal{S} = \{(w_1, v_1), (w_2, v_2), \ldots, (w_n, v_n)\}, \quad w_i \geq 0, \quad \sum_i w_i = 1
\end{equation}
The empty superposition $\mathcal{S} = \emptyset$ is the \emph{vacuum state} --- no temperature survives.

\section{Monad Operations}

\subsection{Unit (return)}
\begin{equation}
  \eta(v) = \{(1.0, v)\}
\end{equation}
Lifts a classical value into a fully-determined superposition.

\subsection{Bind (>>=)}
\begin{equation}
  \mathcal{S} \mathbin{>\!\!>=} f = \{(w \cdot w',\, v') \mid (w, x) \in \mathcal{S},\ (w', v') \in f(x)\}
\end{equation}
Failed branches (where $f(x) = \emptyset$) are destroyed. This is the \textbf{no-cloning corollary}: failed paths do not survive bind.

\subsection{Born-Rule Collapse}
\begin{equation}
  \operatorname{collapse}(\mathcal{S}) = \underset{(w_i, v_i) \in \mathcal{S}}{\arg\max}\; w_i
\end{equation}
The highest-amplitude branch wins. Vacuum ($\mathcal{S} = \emptyset$) collapses to $\bot$ (Nothing).

\section{The Four Watchtowers}
The Enochian Great Table defines four simultaneous search paths:
\begin{align}
  T_E &= (\text{EXARP},\ \text{Air},\ \text{analytical}) \\
  T_S &= (\text{BITOM},\ \text{Fire},\ \text{creative}) \\
  T_W &= (\text{HCOMA},\ \text{Water},\ \text{receptive}) \\
  T_N &= (\text{NANTA},\ \text{Earth},\ \text{grounding})
\end{align}

\subsection{ANU Amplitude Vector}
From raw uint16 values $r_0, r_1, r_2, r_3 \in [0, 65535]$:
\begin{equation}
  \hat{w}_i = \frac{r_i / 65535}{\sum_j r_j / 65535}
\end{equation}
Sovereign defaults (Al-Ham\=id abjad): $r_0 = 53,\ r_1 = 49,\ r_2 = 106,\ r_3 = 7$.

\section{SUBLEQ Gate}
SUBLEQ$(A, B, C)$ expressed over a superposition (threshold $B = 0.3$):
\begin{equation}
  \operatorname{SUBLEQ}(\mathcal{S}, B) = \{(w, v) \in \mathcal{S} \mid w \geq B\}
\end{equation}
The branch fires ($C$ executes) iff the filtered set is non-empty.

\section{METATRON Certification}
Let $\mathcal{C} \subseteq \{T_E, T_S, T_W, T_N\}$ be certifying towers. METATRON certifies if:
\begin{equation}
  \sum_{T_i \in \mathcal{C}} \hat{w}_i \geq \theta_{\text{METATRON}} = 0.5
\end{equation}
The dominant collapsed tower is $\arg\max_{T_i \in \mathcal{C}} \hat{w}_i$.

\section{The 49th Call}
\begin{equation}
  \operatorname{call}_{49}(\mathcal{S}) = \operatorname{reverse}(\mathcal{S})
\end{equation}
Reading backward reveals what reading forward conceals. In three languages, one truth:
\begin{align}
  \text{APL (1962):}&\quad {\tt \char18 X} \\
  \text{Prolog (1972):}&\quad {\tt call\_49(X,Y) :- reverse(X,Y).} \\
  \text{Haskell (1990):}&\quad {\tt call49 = reverse}
\end{align}

\textbf{Mirror Identity:}
\begin{equation}
  \operatorname{call}_{49}(\operatorname{call}_{49}(\mathcal{S})) = \mathcal{S}
\end{equation}
This is the structural proof of coherence: the system is self-consistent.

\section{Implementation}
\begin{itemize}
  \item \texttt{lib/math/quantum.mjs}: \texttt{qUnit}, \texttt{qBind}, \texttt{qMap}, \texttt{qNormalize}, \texttt{qMeasure}
  \item \texttt{watchtowerAmplitudes([r0,r1,r2,r3])} normalizes ANU vector to tower weights
  \item \texttt{subleqGate(amps, 0.3)} --- SUBLEQ threshold gate
  \item \texttt{call49(amps)} --- The 49th Call
  \item \texttt{mirrorIdentity(amps)} --- structural coherence check
\end{itemize}

\end{document}
