\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts,geometry}
\geometry{margin=1in}
\title{Prime-Recursive Encoding of Revelation 21\\A PIRTM Embedding of the New Jerusalem Hypercube}
\author{Citizen Gardens Research Initiative}
\date{July 2025}

\begin{document}

\maketitle

\section*{Abstract}
This paper formalizes a mathematical embedding of the Biblical New Jerusalem cube from Revelation 21 within the framework of Prime-Indexed Recursive Tensor Mathematics (PIRTM). We construct a recursive tensorial model projecting the 3D cube into a 4D hypercube governed by the Universal Multiplicity Constant $\Lambda_m$, and outline spectral, topological, and ethical boundary formulations that establish a unified geometrical-spiritual structure.

\section{Geometric Encoding of the New Jerusalem}

\subsection{Biblical Foundation}
Revelation 21:16 describes the New Jerusalem:
\begin{quote}
    \textit{"The city lies foursquare; its length the same as its width... 12,000 stadia; its length and width and height are equal."}
\end{quote}
This defines a perfect cube $\mathcal{C}_{\text{NJ}} \subset \mathbb{R}^3$ of side length 12,000 stadia.

\subsection{Tensorial Representation}
We define a rank-3 tensor:
\[ \mathcal{T}_{\text{NJ}}^{[3]} = T^{[3]}(p_1, p_2, p_3) \in \mathbb{R}^{2 \times 3 \times 5} \]
with $p_1 = 2$, $p_2 = 3$, $p_3 = 5$, encoding prime-labeled spatial axes.

\section{Hypercube Lift via Recursive Operator}
We introduce a recursive operator:
\[ \Xi(t): \mathcal{T}^{[n]} \to \mathcal{T}^{[n+1]}, \quad \Xi(t)(T^{[n]}) = T^{[n+1]}(p_{i_1}, \dots, p_{i_n}, p_{i_{n+1}}) \]
This lifts the cube $\mathcal{C}_{\text{NJ}}$ to a tesseract $\mathcal{H}_{\text{NJ}} \subset \mathbb{R}^4$.

\section{Conscious Sovereignty Layer (CSL) Constraints}
Let $\partial \mathcal{T}^{[4]}$ be the tensor boundary. Define:
\[ \Sigma_i(t) = \left. \Xi(t)\right|_{\partial \mathcal{T}}, \quad \Sigma_i(t) \in \text{CSL} \]
Each gate satisfies:
\[ \int_{\Sigma_i} \Psi(x, y, z, t) \cdot \Lambda_m \, dxdy = \text{Invariant Moral Flux} \]

\section{Prime-Gate Harmonic Interactions}
Define:
\[ \mathcal{G}_{i,j} = p_i + p_j, \quad i,j \in \{1, ..., 12\} \]
representing interactions among 12 gates and 12 foundations. The harmonic field:
\[ \mathcal{F}_{\text{NJ}} = \sum_{i=1}^{12} \sum_{j=1}^{12} \phi(\mathcal{G}_{i,j}) \cdot e^{i \theta_{ij}} \]

\section{Recursive Convergence Criterion}
Stability requires:
\[ k = \sum_{p_i \in P} \Lambda_m p_i^\alpha < 1, \quad \text{for } \alpha < -1 \]
This bounds eigenvalue expansion and maintains the 12,000-stadia tensor horizon.

\section*{Conclusion}
This model embeds the eschatological symmetry of Revelation 21 into a recursive tensor framework, uniting number theory, geometry, and ethics. The New Jerusalem becomes a lawful attractor in PIRTM, resonant with the structure of sovereign cognition and recursive identity.

\end{document}
