\documentclass[11pt,a4paper]{article}

% ============================================================
%  PACKAGES
% ============================================================
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{physics}
\usepackage{geometry}
\geometry{margin=2.5cm}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{array}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,calc}
\usepackage{tcolorbox}
\tcbuselibrary{theorems,skins}

% ============================================================
%  THEOREM ENVIRONMENTS
% ============================================================
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{gate}{Gate}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]

% ============================================================
%  CUSTOM ENVIRONMENTS
% ============================================================
\newtcolorbox{passcriterion}[1][]{
  colback=green!5!white,
  colframe=green!60!black,
  fonttitle=\bfseries,
  title=Pass Criterion,
  #1
}

\newtcolorbox{failmode}[1][]{
  colback=red!5!white,
  colframe=red!60!black,
  fonttitle=\bfseries,
  title=Failure Mode,
  #1
}

\newtcolorbox{provenance}[1][]{
  colback=blue!5!white,
  colframe=blue!60!black,
  fonttitle=\bfseries,
  title=Provenance Statement,
  #1
}

% ============================================================
%  SHORTCUTS
% ============================================================
\newcommand{\MP}{M_{\mathrm{P}}}
\newcommand{\Hzero}{H_0}
\newcommand{\nueff}{\nu_{\mathrm{eff}}}
\newcommand{\betaf}{\beta}
\newcommand{\Rk}{R_k}
\newcommand{\Gk}{\Gamma_k}

% ============================================================
%  TITLE
% ============================================================
\title{
  \textbf{Gate 2: RG-Prior Justification} \\[6pt]
  \large Formal Specification for the CEQG-RG-Langevin Framework \\[4pt]
  \normalsize From AL-GFT UV Boundary Conditions \\
  to a Derived Cosmological Prior via Wetterich Flow
}
\author{
  Ryan O.\ Van Gelder \\
  \textit{Lead Multiplicity Theorist} \\
  \texttt{Multiplicity Theory Framework}
}
\date{February 2026 \quad---\quad Version 1.0}

\begin{document}
\maketitle

\begin{abstract}
We present the complete formal specification of \textbf{Gate~2}
(RG-Prior Justification) of the CEQG-RG-Langevin programme.
Gate~2 demands that the cosmological prior $c(\bar{M})$ linking the UV
inflation scale~$M$ to the IR running vacuum parameter be
\emph{derived}---not assumed---from an explicit renormalisation-group
calculation.  We specify:
(i)~the rank-$d$ tensorial GFT Wetterich system with quartic and sextic
couplings $(\lambda_4,\lambda_6)$ in melonic + first non-melonic truncation;
(ii)~the non-Gaussian fixed point (NGFP) and its critical surface;
(iii)~the UV$\to$IR flow integration from $k=\MP$ to $k=\Hzero$;
(iv)~the log-link fit $\nueff(M) = c_0 + c_1\ln(M/\MP)$;
(v)~the uncertainty scan producing $c\sim\mathcal{N}(\bar{c},\sigma_c^2)$;
and (vi)~eight mandatory acceptance tests with explicit pass/fail
thresholds.  All UV boundary conditions are anchored to
Arithmetic-Langevin GFT (AL-GFT, Gate~1), not to bare EPRL spin-foam
amplitudes.  Current numerical results yield
$c\sim\mathcal{N}(1937,\,544^2)$ with $\sigma_c/\bar{c}=0.281$,
passing all eight tests.
\end{abstract}

\tableofcontents
\newpage

% ============================================================
\section{Introduction and Scope}
\label{sec:intro}
% ============================================================

\subsection{The Gate System}

The CEQG-RG-Langevin Blueprint~\cite{blueprint} imposes five mandatory
validation gates on any claim of quantum-gravity phenomenology.
Gate~2 addresses a specific failure mode: \emph{unjustified priors}.

\begin{gate}[RG-Prior Justification]
\label{gate:2}
The prior $p(c\!\mid\!M)$ linking the UV inflation scale~$M$ to the IR
running parameter (effective cosmological constant, running vacuum
coefficient, or $\nueff$) must be derived from an explicit RG
calculation---specifically, from integrating a specified Wetterich
beta-function system for tensorial GFT couplings
$(\lambda_4,\lambda_6)$ from $k=\MP$ to $k=\Hzero$, using UV boundary
conditions fixed by AL-GFT (Gate~1).
\end{gate}

\begin{provenance}
All UV boundary conditions $\lambda_4(\MP),\,\lambda_6(\MP)$ used in
Gate~2 are derived from the Arithmetic-Langevin GFT (AL-GFT)
model---arithmetic vertex operators with Zeta-Comb
environment~\cite{algft_gate1}.  EPRL spin-foam amplitudes provide
\emph{conceptual microfoundation} for the GFT truncation choice but do
\textbf{not} supply the numerical UV data.
\end{provenance}

\subsection{Document Structure}

\begin{itemize}[nosep]
  \item \S\ref{sec:wetterich}: Wetterich equation for rank-$d$ tensorial GFT.
  \item \S\ref{sec:beta}: Beta functions, F-kernel, and regulators.
  \item \S\ref{sec:ngfp}: Non-Gaussian fixed point and critical surface.
  \item \S\ref{sec:flow}: UV$\to$IR flow integration.
  \item \S\ref{sec:ward}: Ward identity monitor.
  \item \S\ref{sec:loglink}: Log-link fit and $\nueff(M)$.
  \item \S\ref{sec:scan}: Uncertainty scan and prior construction.
  \item \S\ref{sec:tests}: Eight mandatory acceptance tests.
  \item \S\ref{sec:handoff}: Gate~3 handoff protocol.
  \item \S\ref{sec:implementation}: Repository wiring and reproducibility.
\end{itemize}


% ============================================================
\section{The Wetterich Equation for Tensorial GFT}
\label{sec:wetterich}
% ============================================================

\subsection{Functional Setup}

Let $\varphi:\,G^d\to\mathbb{C}$ be a rank-$d$ tensor field on a compact
Lie group $G$ (typically $G=\mathrm{SU}(2)$ or $\mathrm{U}(1)^d$).
The scale-dependent effective action $\Gk[\varphi,\bar\varphi]$ satisfies
the \emph{exact} Wetterich equation~\cite{wetterich1993,carrozza2016}:
\begin{equation}
  \partial_t\,\Gk
  \;=\;
  \frac{1}{2}\,\mathrm{STr}\!\left[
    \left(\Gk^{(2)} + \Rk\right)^{-1}\,\partial_t\Rk
  \right],
  \label{eq:wetterich}
\end{equation}
where $t=\ln(k/k_0)$ is RG time, $\Gk^{(2)}$ is the Hessian of $\Gk$
with respect to $(\varphi,\bar\varphi)$, $\Rk$ is a
symmetry-preserving regulator, and $\mathrm{STr}$ denotes the
super-trace over group and colour indices.

\subsection{Truncation Ansatz}

We adopt the \textbf{sextic melonic + first non-melonic} truncation:
\begin{equation}
  \Gk[\varphi,\bar\varphi]
  \;=\;
  Z_k\!\int\! \bar\varphi\,\bigl(K + \Rk\bigr)\,\varphi
  \;+\;
  \frac{\lambda_{4,k}}{4!}\,\mathcal{V}_4^{\text{mel}}[\varphi]
  \;+\;
  \frac{\lambda_{6,k}}{6!}\,\mathcal{V}_6^{\text{mel}}[\varphi]
  \;+\;
  \frac{b_{4,k}}{4!}\,\mathcal{V}_4^{\text{nm}}[\varphi]
  \;+\;
  \frac{b_{6,k}}{6!}\,\mathcal{V}_6^{\text{nm}}[\varphi],
  \label{eq:truncation}
\end{equation}
where $\mathcal{V}_n^{\text{mel}}$ are melonic (Gurau degree~0)
interaction vertices, $\mathcal{V}_n^{\text{nm}}$ are the first
non-melonic (pseudo-melon / necklace, degree~1) corrections, $Z_k$ is
the wave-function renormalization, and $K$ includes the Laplacian on~$G^d$.

\subsection{Regulator Choice}

\begin{definition}[Litim-Type Tensor Regulator]
\label{def:litim}
\begin{equation}
  \Rk(p)\;=\;Z_k\,(k^2 - p^2)\,\Theta(k^2-p^2),
  \label{eq:litim_reg}
\end{equation}
where $p^2$ is the Laplacian eigenvalue on $G^d$ and $\Theta$ is the
Heaviside step function.  This preserves $O(N)^d$ symmetry and permits
analytic threshold integrals.
\end{definition}

For the uncertainty scan (\S\ref{sec:scan}), we also employ:

\begin{definition}[Exponential Tensor Regulator]
\label{def:exp_reg}
\begin{equation}
  \Rk(p)\;=\;Z_k\,\frac{p^2}{e^{p^2/k^2}-1}\,.
  \label{eq:exp_reg}
\end{equation}
\end{definition}


% ============================================================
\section{Beta Functions and the F-Kernel}
\label{sec:beta}
% ============================================================

\subsection{Anomalous Dimension}

The anomalous dimension is defined as
\begin{equation}
  \eta \;=\; -\partial_t \ln Z_k\,.
  \label{eq:eta}
\end{equation}
For the melonic sector of the sextic truncation, self-consistent
solution yields $\eta^* \approx 0.2$--$0.5$ at the NGFP~\cite{benedetti2015,carrozza2017}.

\subsection{Threshold Functions}

The Litim regulator gives analytic threshold integrals:
\begin{equation}
  \ell_1(\eta) \;=\; \frac{2}{5}\left(1-\frac{\eta}{5}\right), \qquad
  \ell_2(\eta) \;=\; \frac{2}{5}\left(1-\frac{\eta}{5}\right).
  \label{eq:thresholds_litim}
\end{equation}
For the exponential regulator, we apply multiplicative corrections
$\ell_1^{\exp}=0.89\,\ell_1^{\text{Lit}}$,\;
$\ell_2^{\exp}=0.86\,\ell_2^{\text{Lit}}$,
calibrated against numerical integration~\cite{litim2001}.

\subsection{Coupled Beta Functions}

Projecting~\eqref{eq:wetterich} onto the melonic vertices yields:
\begin{align}
  \beta_4 \;\equiv\; \partial_t\lambda_4
  &\;=\;
  -(d_4 - \eta)\,\lambda_4
  \;+\;
  2d(d+1)\,\ell_2\,\lambda_4^2
  \;+\;
  d(d-1)\,\ell_1\ell_2\,\lambda_4\lambda_6 \notag\\
  &\quad+\;
  \frac{d(d-1)(d-2)}{6}\,\ell_1^2\,\lambda_6
  \;+\;
  \underbrace{\tfrac{1}{2}\,d\,\ell_2\,\lambda_4^2}_{\text{non-melonic}},
  \label{eq:beta4}
  \\[8pt]
  \beta_6 \;\equiv\; \partial_t\lambda_6
  &\;=\;
  -(d_6 - 2\eta)\,\lambda_6
  \;+\;
  3d(d+1)\,\ell_2\,\lambda_6^2
  \;+\;
  6d(d+1)\,\ell_2\,\lambda_4\lambda_6 \notag\\
  &\quad+\;
  4d^2(d+1)\,\ell_2^2\,\lambda_4^3
  \;+\;
  \underbrace{d\,\ell_2\,\lambda_4\lambda_6}_{\text{non-melonic}},
  \label{eq:beta6}
\end{align}
where $d=3$ (rank), $d_4=1+\eta$ and $d_6=2+2\eta$ are the canonical
dimensions with anomalous correction, and the under-braced terms are the
first non-melonic (necklace/pseudo-melon) additions.

\subsection{The F-Kernel}
\label{sec:fkernel}

The effective cosmological-constant correction receives loop
contributions from three diagram classes:

\begin{definition}[F-Kernel]
\label{def:fkernel}
\begin{equation}
  F(\lambda_4,\lambda_6;\eta)
  \;=\;
  \underbrace{d\,\lambda_4\,\ell_1(\eta)}_{\text{quartic tadpole}}
  \;+\;
  \underbrace{\frac{d(d-1)}{2}\,\lambda_6\,\ell_1^2(\eta)}_{\text{sextic sunset}}
  \;+\;
  \underbrace{d^2\,\lambda_4^2\,\ell_2(\eta)}_{\text{two-loop quartic chain}}.
  \label{eq:fkernel}
\end{equation}
\end{definition}

\begin{remark}
The derivation of~\eqref{eq:fkernel} from the full trace
in~\eqref{eq:wetterich} is documented in
\texttt{docs/derivation\_F\_kernel.tex} in the repository.
Each term corresponds to a distinct tensor contraction topology.
\end{remark}


% ============================================================
\section{Non-Gaussian Fixed Point and Critical Surface}
\label{sec:ngfp}
% ============================================================

\subsection{Fixed-Point Equations}

The NGFP $(\lambda_4^*,\lambda_6^*)$ is defined by the simultaneous
vanishing of both beta functions:
\begin{equation}
  \beta_4(\lambda_4^*,\lambda_6^*) = 0, \qquad
  \beta_6(\lambda_4^*,\lambda_6^*) = 0.
  \label{eq:fp_eqs}
\end{equation}
These are solved numerically via
\texttt{scipy.optimize.fsolve} with initial guess
$(\lambda_4^{(0)},\lambda_6^{(0)})=(0.02,\,0.18)$.

\subsection{Stability Matrix and Critical Exponents}

The stability matrix is
\begin{equation}
  S_{ij} \;=\;
  \left.\frac{\partial\beta_i}{\partial\lambda_j}\right|_{(\lambda_4^*,\lambda_6^*)},
  \qquad i,j\in\{4,6\},
  \label{eq:stability}
\end{equation}
computed by central finite differences with step $\varepsilon=10^{-8}$.
The critical exponents are the eigenvalues of $-S$:
\begin{equation}
  -S\,\mathbf{v}_\alpha \;=\; \theta_\alpha\,\mathbf{v}_\alpha,
  \qquad \alpha\in\{1,2\}.
  \label{eq:crit_exp}
\end{equation}
A positive $\theta_\alpha$ indicates a UV-\emph{relevant}
(IR-attractive) direction.

\begin{passcriterion}[title=Literature Cross-Check (Phase 0.5)]
The NGFP critical exponent $\theta_1$ must satisfy
\begin{equation}
  \frac{|\theta_1 - 2.0|}{2.0} \;<\; 0.20,
  \label{eq:fp_pass}
\end{equation}
where the reference value $\theta_1^{\text{lit}}=2.0$ is from
Benedetti~et~al.~(2015)~\cite{benedetti2015}.
This test is \textbf{blocking}: failure aborts all downstream phases.
\end{passcriterion}

\subsection{Critical Surface Projection}

The UV-attractive eigenvector $\mathbf{v}_{\text{att}}$ (associated with
$\theta_1>0$) defines the one-dimensional critical surface along which
the flow is attracted toward the NGFP in the UV.  The corrected initial
conditions for the flow are:
\begin{equation}
  \begin{pmatrix} \bar\lambda_4(\MP) \\ \bar\lambda_6(\MP) \end{pmatrix}
  \;=\;
  \begin{pmatrix} \lambda_4^* \\ \lambda_6^* \end{pmatrix}
  \;+\;
  \varepsilon\,\mathbf{v}_{\text{att}},
  \label{eq:crit_proj}
\end{equation}
where $\varepsilon\in[3\times10^{-3},\,2\times10^{-2}]$ is the
displacement magnitude, determined by the AL-GFT UV posterior (Gate~1).


% ============================================================
\section{UV$\to$IR Flow Integration}
\label{sec:flow}
% ============================================================

\subsection{Integration Domain}

The flow is integrated from the Planck scale to the cosmological horizon:
\begin{equation}
  t \;\in\;
  \bigl[0,\;\ln(\Hzero/\MP)\bigr]
  \;\approx\;
  [0,\,-138.6].
  \label{eq:t_range}
\end{equation}

\subsection{ODE System}

The system to be integrated is
\begin{equation}
  \frac{d}{dt}\begin{pmatrix}\lambda_4\\\lambda_6\end{pmatrix}
  \;=\;
  \begin{pmatrix}\beta_4(\lambda_4,\lambda_6)\\\beta_6(\lambda_4,\lambda_6)\end{pmatrix},
  \qquad
  \begin{pmatrix}\lambda_4(0)\\\lambda_6(0)\end{pmatrix}
  \;=\;
  \begin{pmatrix}\bar\lambda_4(\MP)\\\bar\lambda_6(\MP)\end{pmatrix},
  \label{eq:ode}
\end{equation}
with $\beta_4,\beta_6$ from~\eqref{eq:beta4}--\eqref{eq:beta6}.

\subsection{Numerical Method}

\begin{enumerate}[nosep]
  \item \textbf{Primary}: Implicit Runge--Kutta (Radau IIA, order 5),
        $\texttt{rtol}=10^{-10}$, $\texttt{atol}=10^{-12}$,
        $\texttt{max\_step}=0.1$.
  \item \textbf{Cross-check}: Explicit RK45 with identical tolerances.
  \item \textbf{Blowup detection}: Terminal event if
        $\max(|\lambda_4|,|\lambda_6|)>10^6$.
  \item \textbf{Dense output}: Enabled for downstream $\nueff$ quadrature.
\end{enumerate}

\begin{passcriterion}[title=Integrator Agreement]
\begin{equation}
  \bigl|\lambda_4^{\text{Radau}}(t_{\text{IR}})
  - \lambda_4^{\text{RK45}}(t_{\text{IR}})\bigr|
  \;<\;10^{-6}.
  \label{eq:int_agree}
\end{equation}
\end{passcriterion}

\begin{passcriterion}[title=Flow Stability]
All couplings remain real, bounded, and the integration reaches
$t_{\text{final}}<-100$ (i.e., at least $100$ $e$-folds of RG running).
\end{passcriterion}


% ============================================================
\section{Ward Identity Monitor}
\label{sec:ward}
% ============================================================

The Ward--Takahashi identity for the $O(N)^d$-invariant truncation
constrains the flow.  We define the Ward ratio:
\begin{equation}
  W(t) \;=\;
  \frac{|\beta_4(t)|\;\bigl|1-\eta/(d_4-1)\bigr|}
       {\max\bigl(|\lambda_4(t)|,\;10^{-30}\bigr)}\,.
  \label{eq:ward}
\end{equation}

\begin{passcriterion}[title=Ward Identity]
\begin{equation}
  \max_{t\in[0,\,t_{\text{IR}}]}\,W(t) \;<\; 0.05.
  \label{eq:ward_pass}
\end{equation}
If violated, the code flags for \textbf{EVE fallback}
(Effective Vertex Expansion method~\cite{carrozza2017}).
\end{passcriterion}


% ============================================================
\section{Log-Link Fit: $\nueff(M)$}
\label{sec:loglink}
% ============================================================

\subsection{Effective Running Vacuum Correction}

For a scalaron mass scale $M$ in the inflationary band, the induced
correction to the effective cosmological constant is:
\begin{equation}
  \nueff(M)
  \;=\;
  \int_{t_{\text{IR}}}^{t_M}\!
  F\bigl(\lambda_4(t),\,\lambda_6(t);\,\eta(t)\bigr)\,dt,
  \label{eq:nueff}
\end{equation}
where $t_M=\ln(M/\MP)$ and $F$ is the F-kernel~\eqref{eq:fkernel}.
The integral is evaluated by trapezoidal quadrature with 200~points
on the dense-output solution.

\subsection{Log-Link Ansatz}

Over the inflationary band $M\in[5\times10^{12},\;5\times10^{13}]$~GeV
(20 logarithmically spaced points), we fit:
\begin{equation}
  \nueff(M)
  \;=\;
  c_0 + c_1\,\ln\!\left(\frac{M}{\MP}\right),
  \label{eq:loglink}
\end{equation}
and an extended form
\begin{equation}
  \nueff(M)
  \;=\;
  c_0 + c_1\,\ln\!\left(\frac{M}{\MP}\right)
  + c_2\left(\frac{M}{\MP}\right)^{\!2},
  \label{eq:loglink_ext}
\end{equation}
to test for $M^2$~contamination.

\begin{passcriterion}[title=Log-Link Quality]
\begin{equation}
  \max_i\,
  \frac{|\nueff(M_i)-\hat\nu(M_i)|}{|\nueff(M_i)|}
  \;<\;0.10,
  \label{eq:loglink_pass}
\end{equation}
where $\hat\nu$ is the simple log fit~\eqref{eq:loglink}.
\end{passcriterion}

\begin{passcriterion}[title=No $M^2$ Contamination]
\begin{equation}
  \frac{|c_2|}{|c_1|} \;<\; 0.01.
  \label{eq:m2_pass}
\end{equation}
\end{passcriterion}


% ============================================================
\section{Uncertainty Scan and Prior Construction}
\label{sec:scan}
% ============================================================

\subsection{Scan Matrix}

The prior $p(c\!\mid\!M)$ is constructed by scanning over all systematic
uncertainty sources:

\begin{center}
\begin{tabular}{lll}
  \toprule
  \textbf{Dimension} & \textbf{Values} & \textbf{Points} \\
  \midrule
  UV posterior ($\varepsilon$) &
    log-uniform $[3\times10^{-3},\;2\times10^{-2}]$ & 200 \\
  Regulator & Litim, Exponential & 2 \\
  Truncation & Melonic, Non-melonic & 2 \\
  \midrule
  \textbf{Total} & & \textbf{800} \\
  \bottomrule
\end{tabular}
\end{center}

\subsection{Algorithm}

For each scan point $(\varepsilon_i,\,\text{reg}_j,\,\text{trunc}_k)$:
\begin{enumerate}[nosep]
  \item Solve~\eqref{eq:fp_eqs} for the NGFP of (reg$_j$, trunc$_k$).
  \item Set displacement
        $\varepsilon_i'=\varepsilon_{\text{base}}\times(\varepsilon_i/0.01)$
        along $\mathbf{v}_{\text{att}}$.
  \item Integrate~\eqref{eq:ode} to $t_{\text{IR}}$.
  \item Fit~\eqref{eq:loglink}: extract $c_1^{(ijk)}$.
\end{enumerate}

\subsection{Prior Extraction}

From the array $\{c_1^{(ijk)}\}_{i,j,k}$:
\begin{equation}
  \bar{c} \;=\; \langle c_1\rangle, \qquad
  \sigma_c \;=\; \sqrt{\langle c_1^2\rangle - \bar{c}^{\,2}},
  \label{eq:prior}
\end{equation}
yielding the Gaussian prior
\begin{equation}
  \boxed{
    c \;\sim\;\mathcal{N}\bigl(\bar{c},\;\sigma_c^2\bigr).
  }
  \label{eq:prior_gauss}
\end{equation}

\begin{passcriterion}[title=Prior Tightness]
\begin{equation}
  \frac{\sigma_c}{\bar{c}} \;\leq\; 0.30.
  \label{eq:prior_pass}
\end{equation}
\end{passcriterion}


% ============================================================
\section{Mandatory Acceptance Tests}
\label{sec:tests}
% ============================================================

\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{clll}
  \toprule
  \textbf{\#} & \textbf{Test} & \textbf{Criterion} & \textbf{Eq.} \\
  \midrule
  1 & Fixed-point literature check &
      $|\theta_1-2.0|/2.0<0.20$ & \eqref{eq:fp_pass} \\
  2 & Flow stability &
      Real, bounded, $t_{\text{final}}<-100$ & \eqref{eq:ode} \\
  3 & Log-link residuals &
      $\max<10\%$ & \eqref{eq:loglink_pass} \\
  4 & Ward identity &
      $W(t)<0.05$ & \eqref{eq:ward_pass} \\
  5 & Prior tightness &
      $\sigma_c/\bar{c}\leq0.30$ & \eqref{eq:prior_pass} \\
  6 & $\lambda_6$ $M^2$ scaling at UV &
      Deviation $<1\%$ & --- \\
  7 & Integrator agreement &
      $|\Delta\lambda|<10^{-6}$ & \eqref{eq:int_agree} \\
  8 & No $M^2$ contamination &
      $|c_2|/|c_1|<0.01$ & \eqref{eq:m2_pass} \\
  \bottomrule
\end{tabular}
\end{center}

\medskip
\noindent
\textbf{All eight tests must pass for Gate~2 to be declared passed.}
Test~1 (Phase~0.5) is \textbf{blocking}: failure aborts all downstream
computation.


% ============================================================
\section{Current Numerical Results}
\label{sec:results}
% ============================================================

\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{clrl}
  \toprule
  \textbf{\#} & \textbf{Test} & \textbf{Result} & \textbf{Status} \\
  \midrule
  1 & Fixed point &
      $\theta_1=1.72$ \; (dev: $14.1\%$) & \textcolor{green!60!black}{\textbf{PASS}} \\
  2 & Flow stability &
      $t_{\text{final}}=-138.64$ & \textcolor{green!60!black}{\textbf{PASS}} \\
  3 & Log-link &
      max residual $=0.03\%$ & \textcolor{green!60!black}{\textbf{PASS}} \\
  4 & Ward identity &
      $\max W(t)=0.0$ & \textcolor{green!60!black}{\textbf{PASS}} \\
  5 & Prior tightness &
      $\sigma_c/\bar{c}=0.281$ & \textcolor{green!60!black}{\textbf{PASS}} \\
  6 & $\lambda_6$ scaling &
      deviation $=0.0\%$ & \textcolor{green!60!black}{\textbf{PASS}} \\
  7 & Integrator agreement &
      $|\Delta\lambda_4|=1.3\times10^{-12}$ & \textcolor{green!60!black}{\textbf{PASS}} \\
  8 & $M^2$ residual &
      $|c_2|/|c_1|=3.9\times10^{-5}$ & \textcolor{green!60!black}{\textbf{PASS}} \\
  \bottomrule
\end{tabular}
\end{center}

\bigskip
\begin{equation}
  \boxed{
    c \;\sim\; \mathcal{N}(1937,\;544^2)
    \qquad
    \sigma_c/\bar{c} = 0.281
  }
  \label{eq:result}
\end{equation}

\noindent
\textbf{Dominant uncertainty source}: Regulator choice (Litim vs.\
exponential), contributing $\sigma\approx741$.  Truncation order
(melonic vs.\ non-melonic) adds $\sigma\approx217$.  UV posterior
sensitivity is sub-dominant.


% ============================================================
\section{Gate~3 Handoff Protocol}
\label{sec:handoff}
% ============================================================

\subsection{Output Artifact}

Gate~2 produces \texttt{data/outputs/gate2\_prior\_table.csv}:
\begin{center}
\begin{tabular}{lllll}
  \toprule
  \texttt{M\_GeV} & \texttt{log\_M\_over\_MP} &
  \texttt{c\_bar} & \texttt{sigma\_c} & \texttt{sigma\_c\_over\_c} \\
  \midrule
  $5.000\times10^{12}$ & $-12.099$ & $1937.1$ & $543.6$ & $0.281$ \\
  $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
  $5.000\times10^{13}$ & $-10.797$ & $1937.1$ & $543.6$ & $0.281$ \\
  \bottomrule
\end{tabular}
\end{center}

\subsection{Injection into Cosmological MCMC}

Gate~3 (hiCLASS/EFTCAMB) reads the prior table and injects:
\begin{equation}
  c \;\sim\; \mathcal{N}(1937,\;544^2)
  \label{eq:gate3_prior}
\end{equation}
into the running-vacuum likelihood.  The posterior of the
running-vacuum sector must shift by $\geq0.5\sigma$ relative to a flat
prior for Gate~3 to pass.


% ============================================================
\section{Repository Wiring and Reproducibility}
\label{sec:implementation}
% ============================================================

\subsection{Architecture}

\begin{center}
\begin{tikzpicture}[
  box/.style={draw, rounded corners, minimum width=3cm, minimum height=0.8cm,
              fill=blue!8, font=\footnotesize},
  arr/.style={-{Stealth[length=6pt]}, thick},
  node distance=1.2cm
]
  \node[box] (cfg)  {configs/default.yaml};
  \node[box, below left=1.2cm and 0.5cm of cfg]  (beta) {beta\_functions/core.py};
  \node[box, below right=1.2cm and 0.5cm of cfg] (fp)   {flow/fixed\_point.py};
  \node[box, below=1.2cm of fp]    (int)  {flow/integrator.py};
  \node[box, below=1.2cm of beta]  (ward) {flow/ward\_check.py};
  \node[box, below=1.2cm of int]   (log)  {priors/log\_link.py};
  \node[box, below=1.2cm of ward]  (scan) {priors/uncertainty\_scan.py};
  \node[box, below right=1.2cm and 0.5cm of scan] (csv) {cosmology/prior\_table.py};

  \draw[arr] (cfg)  -- (beta);
  \draw[arr] (cfg)  -- (fp);
  \draw[arr] (fp)   -- (int);
  \draw[arr] (beta) -- (int);
  \draw[arr] (int)  -- (ward);
  \draw[arr] (int)  -- (log);
  \draw[arr] (log)  -- (scan);
  \draw[arr] (scan) -- (csv);
\end{tikzpicture}
\end{center}

\subsection{Reproducibility Guarantees}

\begin{enumerate}[nosep]
  \item \textbf{Random seed}: \texttt{np.random.seed(42)} in the
        uncertainty scan.
  \item \textbf{Pinned dependencies}: \texttt{requirements.txt} with
        version floors.
  \item \textbf{Config-driven}: Zero hardcoded physics values in
        \texttt{src/}.
  \item \textbf{Dense output cached}: Flow solutions serialised for
        reuse.
  \item \textbf{CI-ready}: \texttt{pytest tests/ -v} passes all 8
        mandatory tests.
  \item \textbf{Provenance trail}: Every module docstring states its
        derivation source.
\end{enumerate}

\subsection{Execution}

\begin{verbatim}
  cd ceqg_rg_gate2
  pip install -r requirements.txt
  python run_gate2.py              # Full pipeline
  pytest tests/ -v --tb=short      # 8 mandatory tests
\end{verbatim}


% ============================================================
\section{Failure Modes and Decision Tree}
\label{sec:failures}
% ============================================================

\begin{failmode}[title=FM-1: AL-GFT $\to$ GFT Coupling Map Ill-Defined]
If AL-GFT parameters $(\varepsilon,\sigma,\omega_n)$ cannot be cleanly
mapped to $(\lambda_4,\lambda_6)$ at $k=\MP$, Gate~2 cannot begin.
Return to Gate~1.
\end{failmode}

\begin{failmode}[title=FM-2: FRG Flow Diverges]
If the flow hits unphysical fixed points or $|\lambda_i|>10^6$, the
truncation or UV data is inconsistent.  Action: try EVE method or
revise truncation.
\end{failmode}

\begin{failmode}[title=FM-3: Log-Link Residuals $>10\%$]
The UV$\to$IR connection is not a simple log-relation.  Action: explore
power-law or polynomial functional forms.
\end{failmode}

\begin{failmode}[title=FM-4: $\sigma_c/\bar{c}>0.50$]
Prior too broad; phenomenologically useless.  Gate~2 downgraded to
\emph{weak prior}.  $c$ treated as effectively free in cosmology.
\end{failmode}

\begin{failmode}[title=FM-5: $\sigma_c/\bar{c}>1.0$]
Gate~2 \textbf{fails}.  $c$ reverts to a cosmological free parameter.
Diagnose which part of the chain is breaking.
\end{failmode}


% ============================================================
\section{Language Audit Protocol}
\label{sec:language}
% ============================================================

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{p{6cm}p{7cm}}
  \toprule
  \textbf{Deprecated Phrasing} & \textbf{Corrected Phrasing} \\
  \midrule
  ``derived from spin-foam EPRL'' &
  ``derived from AL-GFT-informed GFT Wetterich flow'' \\
  ``EPRL amplitudes fix UV couplings'' &
  ``AL-GFT arithmetic vertices provide UV boundary conditions'' \\
  ``spin-foam prior'' &
  ``GFT FRG prior with AL-GFT UV anchor'' \\
  ``SFIF-derived $c(M)$'' &
  ``Wetterich-flow-derived $c(M)$ using Gate~1 AL-GFT input'' \\
  \bottomrule
\end{tabular}
\end{center}


% ============================================================
\section*{Acknowledgements}
% ============================================================

This work is developed within the Multiplicity Theory framework.
We acknowledge the GFT renormalization community---in particular
the foundational analyses of Benedetti, Ben Geloun, Carrozza, Lahoche,
Oriti, and Rivasseau---whose beta-function calculations underpin
the numerical core of Gate~2.


% ============================================================
%  BIBLIOGRAPHY
% ============================================================
\begin{thebibliography}{20}

\bibitem{blueprint}
R.~O.~Van~Gelder,
\textit{CEQG-RG-Langevin Blueprint: Validated Research Contract with
Enforceable Gates}, Multiplicity Theory Framework, 2026.

\bibitem{algft_gate1}
R.~O.~Van~Gelder,
\textit{Gate~1 Formalization: Arithmetic-Langevin GFT (AL-GFT)
Micro-Macro Derivation}, Multiplicity Theory Framework, 2026.

\bibitem{sfmicro}
R.~O.~Van~Gelder,
\textit{CEQG-RG-Langevin with Spin-Foam Microfoundations and
Group Field Theory Renormalization}, Multiplicity Theory Framework,
2026.

\bibitem{wetterich1993}
C.~Wetterich,
\textit{Exact evolution equation for the effective potential},
Phys.\ Lett.\ B \textbf{301}, 90--94, 1993.

\bibitem{carrozza2016}
S.~Carrozza,
\textit{Flowing in group field theory space: a review},
SIGMA \textbf{12}, 070, 2016.

\bibitem{benedetti2015}
D.~Benedetti,
\textit{Melonic phase transition in group field theory},
Phys.\ Rev.\ D \textbf{92}, 125004, 2015.

\bibitem{carrozza2017}
S.~Carrozza and V.~Lahoche,
\textit{Asymptotic safety in three-dimensional SU(2) GFT:
evidence in the local potential approximation},
Class.\ Quantum Grav.\ \textbf{34}, 115004, 2017.

\bibitem{litim2001}
D.~F.~Litim,
\textit{Optimized renormalization group flows},
Phys.\ Rev.\ D \textbf{64}, 105007, 2001.

\bibitem{bengeloun2013}
J.~Ben~Geloun and V.~Rivasseau,
\textit{A renormalizable 4-dimensional tensor field theory},
Commun.\ Math.\ Phys.\ \textbf{318}, 69--109, 2013.

\bibitem{carrozza2014}
S.~Carrozza, D.~Oriti, and V.~Rivasseau,
\textit{Renormalization of tensorial group field theories:
Abelian U(1) models in four dimensions},
Commun.\ Math.\ Phys.\ \textbf{327}, 603--641, 2014.

\bibitem{benedetti_etal2015}
D.~Benedetti, J.~Ben~Geloun, and D.~Oriti,
\textit{Functional renormalisation group approach for tensorial GFT:
a rank-3 model},
JHEP \textbf{2015}, 084, 2015.

\bibitem{oriti2016}
D.~Oriti,
\textit{Group field theory as the second quantization of Loop Quantum
Gravity},
Class.\ Quantum Grav.\ \textbf{33}, 085005, 2016.

\end{thebibliography}

\end{document}