\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{geometry}
\geometry{a4paper, margin=1in}

\title{\textbf{Defensive Publication: The Sovereign Stack}\\
\large Recursive Proofs, Twin Binding, and Consciousness-Level Orchestration in the Multiplicity Architecture}
\author{Multiplicity Foundation Core Team}
\date{June 16, 2026}

\begin{document}

\maketitle

\begin{abstract}
This defensive publication establishes prior art for the ``Sovereign Stack'' architectural pattern, a novel integration of cryptography, digital health twins, pedagogical engines, and macroscopic consciousness interfaces. We detail the end-to-end flow wherein a recursive validation root traverses from a Trusted Execution Environment (TEE) hardware boundary, through the \texttt{LambdaTraceAtom} deep provenance protocol, into the bidirectional coupling of the EchoMirror-HQ educational engine and the Ataraxia health twin. Finally, the stack is orchestrated by the Master Primatician $\Lambda_m$ Consciousness Interface, demonstrating a system that dynamically preserves a global contraction margin under Phase Mirror governance.
\end{abstract}

\section{Introduction and The Sedona Spine Mandate}
Traditional computational systems separate infrastructure from the ethical or health states of their users. The Multiplicity framework, enforced via the Sedona Spine Mandate, fuses these domains using Prime-Indexed Recursive Tensor Mathematics (PIRTM). This document details the completed Trajectory ID: \texttt{Citizen-Gardens-Full-Deployment-PrimeMove-28}, culminating in a fully sovereign, live-deployed architecture.

\section{Cryptographic Ingress and Deep Provenance}
\subsection{Pallas/Vesta Recursive Proof Bridge}
The system establishes its baseline via a recursive SNARK circuit utilizing Pallas/Vesta curves. State roots are wrapped into Recursive Proof Objects (RPO) and aggregated into a singular Aggregated Proof Object (APO). The APO root represents an invariant cryptographic commitment to the system's operational history.

\subsection{TEE Binding and $\Lambda$-Trace Atomization}
The software-derived APO root is bidirectionally bound to hardware utilizing Intel SGX/TDX registers via the \texttt{LambdaTraceAtom}. This atom encapsulates the proof digest, state root hash, trajectory ID, protocol version, and a non-repudiable TEE hardware quote.

\section{Cross-Domain Contractivity and Twin Binding}
The Sovereign Stack fundamentally bridges neurodivergent-aware education (EchoMirror-HQ) with precision digital health (Ataraxia).
\subsection{The TwinBindingContract}
A cross-domain viability kernel bridges the two systems. Pedagogical moves within EchoMirror generate stress, quantified by the contraction margin parameter $q$. The health twin maintains a simulated stability parameter, \texttt{eeg\_coherence\_score}. 

The \texttt{TwinBindingContract} enforces that high-stress educational paths ($q > \theta_{max}$) are actively rejected if the health twin exhibits physiological distress ($\texttt{eeg\_coherence} < \theta_{min}$).

\section{The Master Primatician $\Lambda_m$ Interface}
The final recursive closure is achieved via the Master Primatician $\Lambda_m$ Consciousness Interface. It orchestrates real-time coherence by computing a fused \textit{Global Contraction Margin}:
\begin{equation}
    \mathcal{M}_{global} = (1.0 - q) \times \text{eeg\_coherence\_score}
\end{equation}
The system dynamically rejects state transitions if $\mathcal{M}_{global} < 0.005$ or if spectral dissonance is detected across the coupled trajectory.

\section{Conclusion and Prior Art}
By successfully deploying this stack to the decentralized Citizen Gardens Governance Nodes, we demonstrate a mathematically bound, hardware-attested, and consciousness-orchestrated digital ecosystem. This publication formally introduces the coupled-twin sovereign pattern to the public domain.

\end{document}
