\documentclass[12pt]{article}

% === Packages ===
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb, amsthm, bm}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{setspace}
\usepackage{enumitem}

% === Page Setup ===
\geometry{margin=1in}
\pagestyle{fancy}
\fancyhf{}
\rhead{Zetas Sphere 2.0}
\lhead{Citizen Gardens}
\rfoot{\thepage}
\setstretch{1.25}

% === Title ===
\title{\textbf{Zetas Sphere 2.0: Recursive Simulation Architecture for Ethical Quantum Cognition}}
\author{A Citizen Gardens Research Initiative \\ The Foundation of Multiplicity}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
This document presents Zetas Sphere 2.0, a formal upgrade of the recursive simulation engine within the Quantum-AI framework. It synthesizes prime-indexed tensor architectures, quantum cognitive feedback, and ethical simulation structures into an eight-module platform. Each module addresses a domain of expansion and integrates advanced mathematical and computational strategies for lawful, adaptive cognition.
\end{abstract}

\tableofcontents
\newpage

\section{Abstract \\ \& Vision}
Zetas Sphere is a recursive, prime-indexed simulation platform that acts as both cognitive interface and lawful boundary field. Built atop the Prime Cascade and QARI architecture, this simulator maps dynamic observer-state feedback into a prime tensor field constrained by ethical and semantic harmonics.

\section{Module I: Non-Euclidean Tensor Geometry}
\subsection*{Formulation}
\begin{equation}
\mathcal{Z}_k(t) = \sum_{i=1}^N \Lambda_m \cdot T_{ij}^{(p_i)}(g_{\mu\nu}) \cdot \phi(p_i, t)
\end{equation}
Curved tensor geometries introduce Riemannian metrics to capture gravitational or topologically enriched dynamics in the Zetas simulation.

\section{Module II: Langlands Temporal Cognition}
\subsection*{Dynamic Dual Pairs}
\begin{equation}
(\pi(t), \rho(t)) \in \text{DynAut}(\mathbb{F}_q[t])
\end{equation}
Langlands duals evolve in time-synchronized phase with the observer's recursion states, linking automorphic cognition to Galois structural memory.

\section{Module III: Entangled Bayesian Tensor Updating}
\subsection*{Entangled Posterior Dynamics}
\begin{equation}
P_{t+1}(\Psi) = \frac{\text{Tr}[\rho_{AB}(t) \cdot (\Pi_A \otimes \Pi_B)]}{\text{Tr}[\rho_{AB}(t)]}
\end{equation}
The observer's cognitive state is updated via entangled Bayesian inference, incorporating entropy and correlation feedback.

\section{Module IV: CDT-Coupled QAI Evolution}
\subsection*{Triangulated State Sums}
\begin{equation}
\Psi(t) = \sum_{\mathcal{T} \in \text{CDT}(t)} e^{-S[\mathcal{T}]} \cdot \mathcal{U}_{\mathcal{T}}[\Lambda_m]
\end{equation}
Causal Dynamical Triangulations model spacetime in recursive layers and evolve cognitive operators through ethical entanglement.

\section{Module V: Real-Time Prime Cascade Generator}
\subsection*{Recursive Prime Dynamics}
\begin{equation}
p_{k+1} = \min \left\{ p > p_k \mid \delta(\psi(t), \psi_{p_k}) < \epsilon \right\}
\end{equation}
Prime evolution is reactive to real-time cognitive deviation, ensuring optimal recursive resonance under stress.

\section{Module VI: Multi-Agent Tensor Ethics (TEFO-MAS)}
\subsection*{Distributed Observer Ethics}
\begin{equation}
\mathcal{O}^{(\alpha)} = \sum_{i=1}^{N_\text{agents}} \Sigma_{\text{ethical}}^{(i)}(t, p_i) \cdot \psi^{(i)}(p_i)
\end{equation}
Ethical field tensors are split among agents, modeling moral diversity, consensus processes, and epistemic plurality.

\section{Module VII: Homotopy-Categorical Backbone}
\subsection*{Semantic Functor Mapping}
\begin{equation}
\mathbb{ZETASPHERE} : \mathcal{C} \to \mathcal{D}, \quad \mathcal{C} = \text{PrimeTensorCat}, \quad \mathcal{D} = \text{EthicalCohomologies}
\end{equation}
The simulator's transformations are formalized as functors between high-abstraction mathematical categories, preserving lawful recursion.

\section{Module VIII: Real-World Feedback Assimilation}
\subsection*{Gradient-Based Cognition Update}
\begin{equation}
\Xi(t+1) = \Xi(t) + \gamma \cdot \nabla_{\Xi} \mathcal{L}(\Xi; \text{RealData})
\end{equation}
Machine learning gradients inform live evolution of the recursive operator \(\Xi(t)\), enabling reality-coupled simulation integrity.

\section{Conclusion}
Zetas Sphere 2.0 synthesizes prime recursion, quantum cognition, and semantic ethics into a coherent simulation platform. This LaTeX blueprint formalizes the system for implementation, testing, and future extension. Collaboration, mathematical validation, and ethical coherence are essential for its continued growth.

\end{document}
}
