\documentclass[12pt]{article}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{physics}
\usepackage{hyperref}
\geometry{margin=1in}

\title{Prime-Origin Tensor Universe (POTU): A Multiplicity-Theoretic Extension of the Entanglement Past Hypothesis}
\author{Authored by: Ava Fronek \\ \\ Enhanced and Formalized by: Ryan O. Van Gelder \\ \\ Community Research Collective \\ \\ Citizen Gardens \\ The Foundation of Multiplicity}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This paper synthesizes Eddy Keming Chen's Entanglement Past Hypothesis (EPH) with the Multiplicity Framework, constructing a Prime-Origin Tensor Universe (POTU) model that embeds time’s arrow into a recursively decohering, prime-partitioned Hilbert space. We propose novel refinements involving golden ratio decoherence, topological braiding, p-adic fractals, and archetypal cognition. The result is a cosmological formalism that unifies entropy dynamics, number theory, holography, and consciousness.
\end{abstract}
\newpage
\section{Introduction}
The Entanglement Past Hypothesis (EPH) posits that the universe began in a highly ordered, low-entanglement state, explaining the arrow of time through quantum decoherence. We integrate this with the Multiplicity Framework, where primes, Fibonacci sequences, and recursive structures define the evolution of tensor fields. We propose that primes provide a natural basis for Hilbert space factorization, leading to both physical and cognitive symmetry breaking.

\section{Prime-Indexed Tensor Factorization}
We model the universe as a tensor field evolving recursively:
\[
T_{t+1}^{(m,n)} = \sum_{p_i \in \mathbb{P}_N} \Lambda_m \cdot p_i^\alpha \cdot T_t^{(m,n)} + F^{(m,n)}
\]
where $\mathbb{P}_N$ is the set of primes up to $N$, and $\Lambda_m$ modulates recursion.

\subsection{Hilbert Partitioning}
\[
\mathcal{H}_\text{Universe} = \bigotimes_{p \in \mathbb{P}} \mathcal{H}_p
\]
Each prime $p$ labels a subsystem whose entanglement evolves with time, creating the arrow of time through recursive decoherence.

\section{Advanced Refinements}

\subsection{Prime-Driven Spacetime Emergence}
Each prime $p$ generates a holographic screen with area entropy:
\[
S_p \propto \sqrt{p}, \quad ds^2 \sim \sum_p p^{-\alpha} \cdot S_p
\]
Primes naturally encode holographic redundancy via p-adic lattices.

\subsection{Golden Ratio Decoherence}
Replace $p^\alpha$ with $p^{\phi t}$ where $\phi = \frac{1 + \sqrt{5}}{2}$:
\[
\Delta S(t) \sim \sum_{p_i} \log \left( p_i^{\phi t} \cdot \| \psi_{p_i}(t) \| \right)
\]

\subsection{Mythic Archetype Operators}
Assign symbolic operators:
\[
\hat{O}_2 = \text{Yin-Yang}, \quad \hat{O}_3 = \text{Trinity}, \quad \hat{O}_5 = \text{Chaos}
\]
\[
T_{t+1}^{(m,n)} \rightarrow \hat{O}_p \cdot T_t^{(m,n)}
\]

\subsection{Negative Primes for Time Reversal}
\[
\mathcal{H}_\text{Universe} = \bigotimes_{p \in \mathbb{P}} \mathcal{H}_p \otimes \mathcal{H}_{-p}
\]
Models anti-entropic clusters and CPT symmetry.

\subsection{Zeta-Controlled Lambda}
\[
\Lambda_m(t) = \frac{1}{\zeta(1/2 + it)}
\]

\section{Unified Equation}
\[
\boxed{
T_{t+1}^{(m,n)} = \sum_{p \in \mathbb{P}_N} \hat{O}_p \cdot \left( p^{\phi t} \cdot \frac{T_t^{(m,n)}}{\zeta(1/2 + it)} \right) + \mathcal{F}^{(m,n)}( \text{anti-primes} )
}
\]

\section{Topological and Cognitive Extensions}
Time braids entangled strands:
\[
T_{t+1}^{(m,n)} = B_p \cdot T_t^{(m,n)}, \quad B_p \in B_N
\]
Knot invariants encode entropy:
\[
S_{\text{vN}}(t) \propto \log |V_K(q)|
\]
Cognitive emergence modeled by:
\[
\Xi_p(t) = \sum_{n} c_n \cdot F_n \cdot \psi_p(t) \cdot A_p
\]

\section{Conclusion}
We present a framework where time, entropy, and cognition emerge from prime-indexed recursive tensor evolution. This synthesis unifies metaphysical intuition with mathematical formalism, paving the way for future simulations and interdisciplinary cosmogenesis.

\end{document}
"""

# Save to LaTeX file
latex_path = Path("/mnt/data/POTU_EPH_Multiplicity_Model.tex")
latex_path.write_text(latex_content)

latex_path.name
